Classless Java

### 2015-11-22 paper/sections/RelatedWork.tex
related work (in progress)

\subsection{Multiple Inheritance in Object oriented languages}
Many authors have argued in favour or against multiple inheritance.  It provides
expressive power, but it is difficult to model, and implement and can create
programs that are hard to reason about.  These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods,
and the yo-yo problem~\cite{taenzer1989problems}.  To conciliate the need for
expressive power and the need for simplicity, lots of models have been proposed
in the past few years, including mixins~\cite{bracha90mixin} and
traits~\cite{scharli03traits}.  They provide novel programming architecture
models in the OO paradigm.

\begin{enumerate}
\item Mixins allows to name components that can be applied to various classes as
  reusable functionality units. However, they suffer from linearisation: the
  order of mixin application is relevant in often subtitle and undesired
  ways. constraints. This hinders their usability and their ability of resolving
  conflicts: the linearisation (total ordering) of mixin inheritance cannot
  provide a satisfactory resolution in all cases and restricts the flexibility
  of mixin composition.

  To fight those limitations, an algebra of mixin operators is
  introduced~\cite{ancona2002calculus}, but this raised the complexity of the
  approach, especially when constructors and fields are
  considered~\cite{marco09FJigsaw}.

\item Simplifying the mixin algebras approach, traits draw a strong line between
  units of reuse (traits) and object factories (classes) In this model,
  traits~\cite{scharli03traits} units of reusable code, containing only methods
  as reusable functionalities. Thus, no state/state initialization is
  considered.

@@

  Some authors see this as good language design fostering good software
  development by helping programmers to think about the structure of their
  programs.  However, other authors see the need of two concepts and the absence
  of state as drawbacks of this model.

\item C++ and Scala also try to provide solutions to multiple inheritance, but
  both suffer from object initialization problems. Virtual inheritance in C++
  provides another solution to multiple inheritance (especially the diamond
  problem by keeping only one copy of the base class)~\cite{ellis1990annotated},
  however suffers from object initialization problem as pointed out by Malayeri
  et al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
  superclasses, which would potentially cause serious semantic error. Scala
  solution (very similar to linearised mixins, but misleadingly called traits in
  the language) avoids this problem by disallowing constructor parameters,
  causing no ambiguity in cases such as diamond problem.  This approach has
  limited expressiveness, and suffers from all the problems of linearised mixin
  composition.
  Python also offers multiple inheritance via linearised mixins. Indeed python any class is implicitly a mixin, and mixin composition informally expressed as\\*
  \Q@ class A use B,C {...new methods...}@\\*
  can be expressed in python as \\*
  \Q@ class Aux: ...new methods...@\\*
  \Q@ class A(B,C,Aux): pass@
\end{enumerate}

\subsection{Multiple Inheritance in Java}
Since Java 8 default methods are introduced, concrete method implementation are
allowed to be defined (via the \textbf{default} keyword) inside
interfaces. Since Java supports implementation of multiple interfaces (instead
of extension of a single class), the introduction of default methods opens the
gate for various flavours of multiple inheritance in Java, using interfaces.
Former work by Bono.et. al.~\cite{bono14}. provides details on mimicking traits
through interfaces.

>>>>

\subsection{Multiple Inheritance in Object Oriented Languages}
Many authors have argued in favour or against multiple inheritance.  It provides
expressive power, but it is difficult to model and implement, and can create
programs that are hard to reason about.  These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.
%and the yo-yo problem~\cite{taenzer1989problems}.  
To conciliate the need for
expressive power and the need for simplicity, lots of models have been proposed
in the past few years, including mixins~\cite{bracha90mixin} and
traits~\cite{scharli03traits}.  They provide novel programming architecture
models in the OO paradigm.

\begin{enumerate}
\item Mixins allows to name components that can be applied to various classes as
  reusable functionality units. However, they suffer from linearisation: the
  order of mixin application is relevant in often subtitle and undesired
  ways. constraints. This hinders their usability and their ability of resolving
  conflicts: the linearisation (total ordering) of mixin inheritance cannot
  provide a satisfactory resolution in all cases and restricts the flexibility
  of mixin composition.  To fight those limitations, an algebra of mixin
  operators is introduced~\cite{ancona2002calculus}, but this raised the
  complexity of the approach, especially when constructors and fields are
  considered~\cite{marco09FJigsaw}.  Our approach does not have the
  linearisation problem, in that the semantics of Java \textbf{extends} clause
  is unordered and symmetric.
\item Simplifying the mixin algebras approach, traits draw a strong line between
  units of reuse (traits) and object factories (classes) In this model,
  traits~\cite{scharli03traits} units of reusable code, containing only methods
  as reusable functionalities. Thus, no state/state initialization is
  considered.

@@

  Some authors see this as good language design fostering good software
  development by helping programmers to think about the structure of their
  programs.  However, other authors see the need of two concepts and the absence
  of state as drawbacks of this model. Our approach takes interfaces as units of
  reuse, and meanwhile generates factory method for instantiation.

\item C++ and Scala also try to provide solutions to multiple inheritance, but
  both suffer from object initialization problems. Virtual inheritance in C++
  provides another solution to multiple inheritance (especially the diamond
  problem by keeping only one copy of the base class)~\cite{ellis1990annotated},
  however suffers from object initialization problem as pointed out by Malayeri
  et al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
  superclasses, which would potentially cause serious semantic error. Scala
  solution (very similar to linearised mixins, but misleadingly called traits in
  the language) avoids this problem by disallowing constructor parameters,
  causing no ambiguity in cases such as diamond problem.  This approach has
  limited expressiveness, and suffers from all the problems of linearised mixin
  composition.
  Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
  \Q@ class A use B,C {...new methods...}@\\*
  can be expressed in python as \\*
  \Q@ class Aux: ...new methods...@\\*
  \Q@ class A(B,C,Aux): pass@
  Our approach not only does not involve linearised mixin problem, but also
  support proper constructor mechanism.
\end{enumerate}

\subsection{Multiple Inheritance in Java}
Since Java 8 default methods are introduced, concrete method implementation are
allowed to be defined (via the \textbf{default} keyword) inside
interfaces. 
% Since Java supports implementation of multiple interfaces (instead
% of extension of a single class), 
The introduction of default methods opens the
gate for various flavours of multiple inheritance in Java, using interfaces.
Former work by Bono.et. al.~\cite{bono14}. provides details on mimicking traits
through interfaces.
